A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-06-27 multiple.tex
Multi-ex

We return again to our example in \cref{sec:intuition} and start by performing
the steps of \cref{sec:single:example} shown in \cref{fig:derivation:single},
since the steps to propagate \cref{ex:string-constraints} across the automaton
$A'$ still applies for the product. We also reuse the definition of $\Map$ as
the simple projection to extract the counter increments, so in this case,
$\Map(\Transition) = \left[l_{B+}\right]$ for every transition~$\Transition$
since we are only counting the length.

We reuse the transition variable associations of our previous
\cref{sec:single:example}, from \cref{eq:single-filter}, but add the definitions
for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6\\
        \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
        \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B})= \TransitionVar_8 \\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}\\
        \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}\\
        \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
        \FlowEq(B, \Filter) =  \begin{aligned}
            1 + \TransitionVar_{12}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

In the derivation tree of \cref{fig:derivation:multi} we start as before, but
with the product version of the $\Image$ predicate. The only possible rule here
is $\ExpandM$, which we use to add the corresponding constraints on each
automata of the product as we would have had in the single-automaton version.

We perform algebraic reasoning on those equations, along with the constraint on
$\MonoidElement$ to determine bounds on transition variables that will enable us
to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
finished doing so, we use \Materialise{} to compute the product of the now
filtered automata. Further algebraic reasoning allows us to remove the final
instance of $\Connected$. We then perform \Expand{} to get rid of the final
$\Image$ instance, and are immediately able to $\Subsume$ the resulting
$\Connected{}$ predicate. This, again, leaves us with a set of linear
inequalities at the leaf that we can use to obtain a model of our final
$\MonoidElement$ values (now a vector). This time we can read out aca, or
$\VectorLiteral{2, 0, 1}$.

\begin{figure}
  \centering
\begin{prooftree}
  % LEFT TREE  ->
  \hypo{l}% here x_12 is zero
  % RIGHT TREE ->
  \hypo{r} % here x_12 is one
  \infer2[\Split{} $\TransitionVar_{12} = 0$]{\text{FINAL STEP OF STUMP}}
  \infer1[\EquationReasoning: probably something good]{
    \begin{aligned}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land \TransitionVar_2 = l_A - 3\\
      \land 1 + \TransitionVar_{12}   = \TransitionVar_{11} + \TransitionVar_{12}\\
      \land \TransitionVar_{11} = \TransitionVar_6\\
      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}\\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{aligned}
  }
  \infer1[\EquationReasoning: expand $\FlowEq$]{
    \begin{aligned}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land \TransitionVar_2 = l_A - 3\\
      \land \FlowEq(B, \Filter) \\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{aligned}
  }
  \infer1[Repeat \cref{fig:derivation:single}, $l_A = l_B$.]{
    \begin{aligned}
      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land i = \TransitionVar_5\\
      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k\\
      \land n = \TransitionVar_4 + \TransitionVar_1\\  
      \land \FlowEq(B, \Filter) \\
      \land \Connected(A', \Filter) \\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) 
      \land l_A = l_B\\
      \land l_A - i - n > 0\\
      \land 0 < i < n
    \end{aligned}
  }
  \infer1[\ExpandM]{
    \Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
    \land l_A = l_B % Implied constraint
    \land l_A - i - n > 0
    \land 0 < i < n
    }
\end{prooftree}
\caption{A derivation for \Calculus{} on the Parikh image strings for the
constraints of \cref{ex:string-constraints}. Note the constant propagation
for~$n$!}\label{fig:derivation:multi}
\end{figure}

>>>>

We return again to our example in \cref{sec:intuition}.We also reuse the
definition of $\Map$ as the simple projection to extract the counter increments,
so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
transition~$\Transition$ since we are only counting the length. The reader is
strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
\cref{sec:intuition} while going through this example! In particular, note that
the transition variables of the figures correspond to the ones used here.

The transition variables of \cref{sec:single:example} recur
\cref{eq:single-filter}, but we add the definitions for $B$ shown in
\cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6\\
        \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
        \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B})= \TransitionVar_8 \\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}\\
        \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}\\
        \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
        \FlowEq(B, \Filter) =  \begin{aligned}
            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

The only possible rule at the start is $\ExpandM$, which we use to add the
corresponding constraints on each automata of the product as we would have had
in the single-automaton version.

We then continue by repeating the steps of the steps of
\cref{sec:single:example} shown in \cref{fig:derivation:single}, since the steps
to propagate \cref{ex:string-constraints} across the automaton $A'$ still apply
for the product. Note that this fixes the values of all transitions but one as
constants.

\Fudge{description here!}

\begin{figure}
  \centering
\begin{prooftree}
  {% % LEFT TREE: here x_12 is zero
  \hypo{PROOF CLOSED}
  \infer1[\Propagate]{WHAT HAPPENS WHEN X_12 IS ZERO}
  }% END LEFT TREE
  {% RIGHT TREE % here x_12 is one
  \hypo{r} 
  }% END RIGHT TREE
  \infer2[\Split{} $\TransitionVar_{12} = 0$]{
  \begin{aligned}
    \TransitionVar_7 = l_A - \TransitionVar_9 - 2\TransitionVar_{11}-2\TransitionVar_{12}\\
    \land \TransitionVar_2 = l_A - 3\\
    \land \TransitionVar_{8} = \TransitionVar_{11} + \TransitionVar_{12} - 1\\
    \land \TransitionVar_6 = \TransitionVar_{11}\\
    \land \TransitionVar_{10} = 1 - \TransitionVar_{11}\\
    \land \Connected(B, \Filter) \\
    \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
  \end{aligned}
  }
  \infer1[Expand $\sum$, \EquationReasoning] {
    \begin{aligned}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land \TransitionVar_2 = l_A - 3\\
      \land 1 + \TransitionVar_{8}   = \TransitionVar_{11} + \TransitionVar_{12}\\
      \land \TransitionVar_{11} = \TransitionVar_6\\
      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}\\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{aligned}
  }
  \infer1[\EquationReasoning: expand $\FlowEq$]{
    \begin{aligned}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land \TransitionVar_2 = l_A - 3\\
      \land \FlowEq(B, \Filter) \\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{aligned}
  }
  \infer1[Repeat \cref{fig:derivation:single}, $l_A = l_B$.]{
    \begin{aligned}
      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land i = \TransitionVar_5\\
      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k\\
      \land n = \TransitionVar_4 + \TransitionVar_1\\  
      \land \FlowEq(B, \Filter) \\
      \land \Connected(A', \Filter) \\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) 
      \land l_A = l_B\\
      \land l_A - i - n > 0\\
      \land 0 < i < n
    \end{aligned}
  }
  \infer1[\ExpandM]{
    \Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
    \land l_A = l_B % Implied constraint
    \land l_A - i - n > 0
    \land 0 < i < n
    }
\end{prooftree}
\caption{A derivation for \Calculus{} on the Parikh image strings for the
constraints of \cref{ex:string-constraints}. Note the constant propagation
for~$n$!}\label{fig:derivation:multi}
\end{figure}
